---
title: 'Reproducible Research: Peer Assessment 1'
output: html_document
keep_md: true
---


## Loading and preprocessing the data

```{r}
activity<-read.csv("activity.csv")
activity$date<-as.Date(activity$date,format="%Y-%m-%d")
```

## What is mean total number of steps taken per day?

##### 1.Loading libraries.
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
```

##### 2.Calculation of the total number of steps taken per day.
```{r}
total_steps_perday<-activity%>%group_by(date)%>%
        summarize(total_steps = sum(steps, na.rm = TRUE))
```

##### 3.Making a histogram of the total number of steps taken each day.
```{r,message=FALSE, warning=FALSE}
ggplot(total_steps_perday, aes(x =total_steps )) +                          
        geom_histogram() +
        labs( x = "Steps", y="Frequency")+
        ggtitle("Total number of steps taken per day")
```

##### 4.Finding mean and median of steps taken per day.
```{r}
mean(total_steps_perday$total_steps, na.rm = TRUE)
median(total_steps_perday$total_steps, na.rm = TRUE)
```


## What is the average daily activity pattern?

##### 1.Calculate average steps grouped by the interval.
```{r}
avg_steps_5min<-activity%>%group_by(interval)%>%
        summarize(avg_steps = mean(steps, na.rm = TRUE))
```

##### 2.Making a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
```{r}
ggplot(avg_steps_5min, aes(x=interval, y=avg_steps)) +
        geom_line(size = I(1.2)) + 
        xlab("5-minute interval")+
        ggtitle("Average number of steps taken")
```

##### 3.Finding which 5-minute interval contains the maximum number of steps.
```{r}
pos_max <- which.max(avg_steps_5min$avg_steps)
avg_steps_5min[pos_max,1]
```

## Imputing missing values

##### 1.The total number of missing values in the dataset
```{r}
sum(is.na(activity$steps))
```

##### 2.Filling in all of the missing values with the mean for that 5-minute interval and creating a new dataset with no missing values.
```{r}
activity1 <- activity %>%
        left_join(activity%>%
                          group_by(interval) %>%
                          summarise(mean_steps = mean(steps, na.rm = TRUE)), 
                  by = c("interval"))%>%
        mutate(steps = ifelse(is.na(steps), mean_steps, steps))%>%
        select(-mean_steps)
```

##### 3.A histogram of the total number of steps taken each day.
```{r, message=FALSE, warning=FALSE}
total_steps_perday1<-activity1%>%group_by(date)%>%
        summarize(total_steps = sum(steps, na.rm = TRUE))

ggplot(total_steps_perday1, aes(x =total_steps )) +                          
        geom_histogram() +
        labs( x = "Steps", y="Frequency")+
        ggtitle("Total number of steps taken per day")
```

##### 4.Calculating mean and median of steps taken per day.
```{r}
mean(total_steps_perday1$total_steps, na.rm = TRUE)
median(total_steps_perday1$total_steps, na.rm = TRUE)
```


## Are there differences in activity patterns between weekdays and weekends?

##### 1.Creating a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r,message=FALSE, warning=FALSE}
Sys.setlocale(category = "LC_ALL", locale = "english")
activity$weekday<-weekdays(activity$date)
activity$weekday[activity$weekday%in%c("Saturday", "Sunday")]<-"weekend"
activity$weekday[activity$weekday!="weekend"]<-"weekday"
activity$weekday<-factor(activity$weekday)
```

##### 2.Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
avg_steps_week<-activity%>%group_by(interval, weekday)%>%
        summarize(avg_steps = mean(steps, na.rm = TRUE))


ggplot(avg_steps_week, aes(interval, avg_steps, colour = factor(weekday))) + 
        geom_line(size = I(1), show.legend = FALSE)+
        facet_grid(rows = vars(weekday))+
        ylab("Number of steps")
```

